---
title: "Differentiating Tape for Trace Evidence"
author: "Yingmai Chen, Maysen Pagan, Chang Shi, Yan Wang"
format: html
editor: visual
number-sections: true
---

# Project Background

# Objectives

# Exploratory Data Analysis

This part provides a concise summary of the data cleaning and organization steps for the four datasets: `width.csv`, `thickness.csv`, `stereo_phys.csv`, and `plm.csv`.

## width.csv

1.  **Data Loading**: The dataset was loaded into the variable `width`.
2.  **Column Removal**: Averages and standard deviation columns were deleted.
3.  **Text Cleaning**: "Tape" was removed from the `X` column.
4.  **Data Type Conversion**: Values in columns 1-10 were converted to numeric.
5.  **Row Removal**: Rows containing NAs and "tape 7" were removed.
6.  **Renaming Columns**: Columns were renamed to a specific format.
7.  **Data Transformation**: The dataset was reshaped to a longer format.

## thickness.csv

1.  **Data Loading**: The dataset was loaded into the variable `thickness`.
2.  **Column Removal**: Averages and standard deviation columns were deleted.
3.  **Text Cleaning**: "Tape" was removed from the `X` column.
4.  **Data Type Conversion**: Values in columns 1-10 were converted to numeric.
5.  **Row Removal**: Rows containing NAs and "tape 7" were removed.
6.  **Renaming Columns**: Columns were renamed to a specific format.
7.  **Data Splitting**: Data was split into `thickness_ba` and `thickness_backing`.
8.  **Data Transformation**: Both subsets were reshaped to a longer format.
9.  **Data Merging**: Merged `width`, `thickness_ba`, and `thickness_backing` into a single dataset named `data`.

## stereo_phys.csv

1.  **Data Loading**: The dataset was loaded into the variable `stereo`.
2.  **Column Removal**: Columns `X.4` to `X.9` were removed.
3.  **Text Cleaning**: "Tape" was removed from the `X` column and the `-` value in the `X.3` column was replaced with "no color".
4.  **Renaming Columns**: Columns were renamed to descriptive names.
5.  **Data Merging**: Physical data from `stereo` was merged with the `data` dataset.

## plm.csv

1.  **Data Loading**: The dataset was loaded into the variable `plm`.
2.  **Column Removal**: Columns `X.7` to `X.12` were removed.
3.  **Row Removal**: Rows corresponding to "tape 7" and the last two rows were removed.
4.  **Renaming Columns**: Columns were renamed to descriptive names.
5.  **Text Cleaning**: The `-` value in the `plm_color` column was replaced with "no color".
6.  **Data Transformation**: The `BOPP` column was converted to binary values.
7.  **Data Merging**: Physical data from `plm` was merged with the `data` dataset.

## Encoding Data

1.  **Encoding Categorical Variables**: Several columns were encoded into numeric values.
2.  **Data Type Conversion**: Converted several categorical columns to numeric.

# Models

The method used to analyze the dissimilarities between each of the 22 tapes involves visualizing a heatmap with hierarchical clusterings of the tapes. These clusterings are determined by the "distances" between each tape.

## Tape Distances {#sec-distances}

Defining a distance between tapes provides a measure of dissimilarity between each tape. Variables for each of the tapes included three numerical variables (width, thickness of backing, and thickness of backing and adhesive) as well as 12 categorical variables. Not all of the variables are numeric and as a result, the Gower's distance is calculated between two tapes which accounts for both numerical and categorical variables. Before taking the Gower's distance between each tape, the categorical variables were first converted to numeric values by encoding each level of the variable 1 to the total number of levels of that variable. For example, the categorical variable `Texture` has three levels: "Rough", "Smooth", "Slightly Rough". "Rough" was coded as 1, "Smooth" was coded as 2, and "Slightly Rough" was coded as 3. Each number was then divided by the total number of levels to obtain normalized variables from 0 to 1. In the example, "Rough" now has the numerical value 0.33, "Smooth" now has the numerical value 0.67, and "Slightly Rough" now has the numerical value 1.

Once all categorical variables were coded numerically, Gower's distance calculates a matrix of dissimilarities for each of the $\frac{n(n-1)}{2} = \frac{22(22-1)}{2} = 231$ pairs of tapes. For the three numerical variables, a range-normalized Manhattan distance is calculated: $$|\frac{T_{ik} - T_{jk}}{\text{range}_k}|$$ where $i, j = 1, 2, \dots, 22$ and $k = 1, 2, 3$. For the remaining 12 categorical variables, the distance is calculated with the following formula: $$I(T_{im} \neq T_{jm})$$ where $i, j = 1, 2, \dots, 22$ and $m = 4, 5, \dots, 15$. This indicator will equal 1 if $T_{im} \neq T_{jm}$ and 0 if $T_{im} \neq T_{jm}$. Gower's distance between two tapes is calculated by summing these variable distances and dividing by 15, the number of variables.

## Heatmap

Once we have the pairwise distances between tapes from the data, we can use a heatmap and multidimensional scaling to visually identify similarities and dissimilarities between tapes. In a heatmap cells are color-coded to quickly compare one row or tape to another. The heatmap also allows one to aggregate the rows in clusters based on the distances calculated from the previous section. The following figure is the heatmap generated for the 22 tapes using Gower's distance. 

```{r}
#| echo: false
#| warning: false

library(pheatmap)
library(ggpubr)
library(factoextra)
library(fmsb)
```

```{r}
#| echo: false
#| warning: false

library(tidyverse)
#data cleaning width
width <- read.csv("width.csv", header = TRUE)
#getting rid of averages and standard deviation columns
width <- width %>% select(-c(5:6, 10:11, 15:20))
#remove "Tape" from X column
width$X <- str_replace(width$X, "Tape ", "")
#make values numeric
width <- data.frame(sapply(subset(width, select = c(1:10)), as.numeric))
#remove row of NAs and remove tape 7 (paper tape)
width <- width[-c(1, 8),]
#add column names
colnames(width) <- c("tape", rep("1", 3), rep("2", 3), rep("3", 3))
#make tape column character
width$tape <- as.factor(width$tape)
#pivot longer
width <- width %>% pivot_longer(!tape, names_to = "site", values_to = "width")


#data cleaning thickness
thickness <- read.csv("thickness.csv", header = TRUE)
#getting rid of averages and standard deviation columns
thickness <- thickness %>% select(-c(5:6, 10:11, 15:19))
#remove "Tape" from X column
thickness$X <- str_replace(thickness$X, "Tape ", "")
#make values numeric
thickness <- data.frame(sapply(subset(thickness, select = c(1:10)), as.numeric))
#remove row of NAs 
thickness <- thickness[-c(1),]
#remove all tape 7 (paper tape)
thickness <- thickness %>% filter(X != 7)
#add column names
colnames(thickness) <- c("tape", rep("1", 3), rep("2", 3), rep("3", 3))
#make tape column character
thickness$tape <- as.factor(thickness$tape)
#thickness of backing and adhesive
thickness_ba <- thickness[c(1:22),]
#thickness of backing
thickness_backing <- thickness[c(23:44),]
#pivot longer
thickness_ba <- thickness_ba %>% pivot_longer(!tape, names_to = "site", values_to = "thickness_backing_adhesive")
thickness_backing<- thickness_backing %>% pivot_longer(!tape, names_to = "site", values_to = "thickness_backing")

#merge
data <- cbind(width, thickness_ba$thickness_backing_adhesive, thickness_backing$thickness_backing)
colnames(data)[4] <- "thickness_backing_adhesive"
colnames(data)[5] <- "thickness_backing"

#get overall average width and thickness for each tape
data <- data %>% group_by(tape) %>% 
  summarize(avg_width = mean(width), 
            avg_thick_back_ad = mean(thickness_backing_adhesive), 
            avg_thick_back = mean(thickness_backing))

```

```{r}
#| echo: false
#| warning: false


#adding stereomicroscope physical data
stereo <- read.csv("stereo_phys.csv")
#remove X.4, X.9
stereo <- stereo %>% select(-c(X.4:X.9))
#remove "Tape" from X column
stereo$X <- str_replace(stereo$X, "Tape ", "")
#remove tape 7
stereo <- stereo[-c(1,8),]
#replace "-" with "no color"
stereo$X.3 <- str_replace(stereo$X.3, "-", "no color")
#new column names
colnames(stereo) <- c("tape", "surface_color", "surface_feature", "wide_narrow", "stereo_color")
#remove spaces in names
stereo$surface_color <- str_replace(stereo$surface_color, "Matte ", "Matte")
stereo$surface_color <- str_replace(stereo$surface_color, "Clear ", "Clear")
#add this physical data to other data set
data <- cbind(data, stereo[,2:5])
```

```{r}
#| echo: false
#| warning: false


#adding plm physical characteristics
plm <- read.csv("plm.csv")
#remove X.7, X.12
plm <- plm %>% select(-c(X.7:X.12))
#remove tape 7
plm <- plm[-c(1,8),]
#new column names
colnames(plm) <- c("tape", "texture", "bubbles", "particle", "edge_texture", "brightness", "plm_color", "BOPP", "extinction")
#replace "-" with "no color"
plm$plm_color <- str_replace(plm$plm_color, "-", "no color")
#remove bottom two rows
plm <- plm[-c(23, 24),]
#BOPP make binary
plm$BOPP <- ifelse(plm$BOPP == "-", 0, 1)
#add this physical data to other data set
data <- cbind(data, plm[,2:9])
```

```{r}
#| echo: false
#| warning: false


#coding all categorical variables
#surface color
data$surface_color <- ifelse(data$surface_color == "Clear", 1, data$surface_color)
data$surface_color <- ifelse(data$surface_color == "Matte", 2, data$surface_color)
data$surface_color <- ifelse(data$surface_color == "Slightly Matte", 3, data$surface_color)
data$surface_color <- ifelse(data$surface_color == "Off-white Matte", 4, data$surface_color)

#surface feature
data$surface_feature <- ifelse(data$surface_feature == "Rough ", 1, data$surface_feature)
data$surface_feature <- ifelse(data$surface_feature == "Circular bubbles", 2, data$surface_feature)
data$surface_feature <- ifelse(data$surface_feature == "Irregular large bubble (water-like)", 3, data$surface_feature)
data$surface_feature <- ifelse(data$surface_feature == "Slightly Rough", 4, data$surface_feature)
data$surface_feature <- ifelse(data$surface_feature == "Large and small circular bubbles", 5, data$surface_feature)
data$surface_feature <- ifelse(data$surface_feature == "Many grooves and bubbles ", 6, data$surface_feature)
data$surface_feature <- ifelse(data$surface_feature == "Minimal surface features ", 7, data$surface_feature)
data$surface_feature <- ifelse(data$surface_feature == "Many grooves ", 8, data$surface_feature)
data$surface_feature <- ifelse(data$surface_feature == "Some small bubbles", 9, data$surface_feature)
data$surface_feature <- ifelse(data$surface_feature == "Large bubble and Irregular bubbles", 10, data$surface_feature)
data$surface_feature <- ifelse(data$surface_feature == "Grooves (water-like)", 11, data$surface_feature)

#wide/narrow
data$wide_narrow <- ifelse(data$wide_narrow == "Wide ", 1, 0)

#stereomicroscope color
data$stereo_color <- ifelse(data$stereo_color == "Pink ", "Pink", data$stereo_color)
data$stereo_color <- ifelse(data$stereo_color == "no color", 1, data$stereo_color)
data$stereo_color <- ifelse(data$stereo_color == "Pink/Green ", 2, data$stereo_color)
data$stereo_color <- ifelse(data$stereo_color == "Purple/Green ", 3, data$stereo_color)
data$stereo_color <- ifelse(data$stereo_color == "Brown ", 4, data$stereo_color)
data$stereo_color <- ifelse(data$stereo_color == "Pink/Green/Yellow/Purple/Blue", 5, data$stereo_color)
data$stereo_color <- ifelse(data$stereo_color == "Slightly Pink ", 6, data$stereo_color)
data$stereo_color <- ifelse(data$stereo_color == "Pink", 7, data$stereo_color)
data$stereo_color <- ifelse(data$stereo_color == "Blue/Pink ", 8, data$stereo_color)
data$stereo_color <- ifelse(data$stereo_color == "Pink/Red/Orange", 9, data$stereo_color)
data$stereo_color <- ifelse(data$stereo_color == "Purple/Blue", 10, data$stereo_color)
data$stereo_color <- ifelse(data$stereo_color == "High Pink ", 11, data$stereo_color)

#texture
data$texture <- ifelse(data$texture == "Rough ", 1, data$texture)
data$texture <- ifelse(data$texture == "Smooth ", 2, data$texture)
data$texture <- ifelse(data$texture == "Slightly Rough ", 3, data$texture)

#bubbles
data$bubbles <- ifelse(data$bubbles == "Y", 1, 0)

#particle
data$particle <- ifelse(data$particle == "Y", 1, 0)

#edge texture
data$edge_texture <- ifelse(data$edge_texture == "Smooth ", 1, 0)

#brightness
data$brightness <- ifelse(data$brightness == "Low Dark ", 1, data$brightness)
data$brightness <- ifelse(data$brightness == "Bright ", 2, data$brightness)
data$brightness <- ifelse(data$brightness == "Dark ", 3, data$brightness)
data$brightness <- ifelse(data$brightness == "Low Bright ", 4, data$brightness)
data$brightness <- ifelse(data$brightness == "High Bright ", 5, data$brightness)
data$brightness <- ifelse(data$brightness == "Bright", 6, data$brightness)

#plm color
data$plm_color <- ifelse(data$plm_color == "no color", 1, data$plm_color)
data$plm_color <- ifelse(data$plm_color == "Green ", 2, data$plm_color)
data$plm_color <- ifelse(data$plm_color == "Slightly green ", 3, data$plm_color)
data$plm_color <- ifelse(data$plm_color == "Pink/Yellow", 4, data$plm_color)
data$plm_color <- ifelse(data$plm_color == "Pink/Yellow/Bright Blue/Turquoise", 5, data$plm_color)
data$plm_color <- ifelse(data$plm_color == "Light Green ", 6, data$plm_color)
data$plm_color <- ifelse(data$plm_color == "Yellow/Gold", 7, data$plm_color)
data$plm_color <- ifelse(data$plm_color == "Yellow/Green", 8, data$plm_color)
data$plm_color <- ifelse(data$plm_color == "Orange", 9, data$plm_color)
data$plm_color <- ifelse(data$plm_color == "Yellow", 10, data$plm_color)

#extinction
data$extinction <- ifelse(data$extinction == "Parallel ", "Parallel", data$extinction)
data$extinction <- ifelse(data$extinction == "Oblique ", "Oblique", data$extinction)
data$extinction <- ifelse(data$extinction == "Symmetrical", "Symmetric ", data$extinction)
data$extinction <- ifelse(data$extinction == "Parallel", 1, data$extinction)
data$extinction <- ifelse(data$extinction == "Symmetric ", 2, data$extinction)
data$extinction <- ifelse(data$extinction == "Oblique", 3, data$extinction)
```

```{r}
#| echo: false
#| warning: false


#convert categorical to numiercal
data$surface_color <- as.numeric(data$surface_color)
data$surface_feature <- as.numeric(data$surface_feature)
data$stereo_color <- as.numeric(data$stereo_color)
data$texture <- as.numeric(data$texture)
data$brightness <- as.numeric(data$brightness)
data$plm_color <- as.numeric(data$plm_color)
data$extinction <- as.numeric(data$extinction)

#normalize all variables
data$avg_width <- data$avg_width/max(data$avg_width)
data$avg_thick_back_ad <- data$avg_thick_back_ad/max(data$avg_thick_back_ad)
data$avg_thick_back <- data$avg_thick_back/max(data$avg_thick_back)
data$surface_color <- data$surface_color/max(data$surface_color)
data$surface_feature <- data$surface_feature/max(data$surface_feature)
data$stereo_color <- data$stereo_color/max(data$stereo_color)
data$texture <- data$texture/max(data$texture)
data$brightness <- data$brightness/max(data$brightness)
data$plm_color <- data$plm_color/max(data$plm_color)
data$extinction <- data$extinction/max(data$extinction)
```

```{r, out.width="200%"}
#| label: fig-heatmap
#| fig-cap: "Clustered Heatmap of 22 Different Tapes. Tape 7 is omitted"
#| echo: false
#| warning: false


#pheatmap
library(philentropy)
row.names(data) <- c(seq(1, 6), seq(8, 23))
library(cluster)
gower_samples <- as.dist(as.matrix(daisy(data[,2:16], metric = "gower")))
gower_features <- as.dist(as.matrix(daisy(t(data[,2:16]), metric = "gower")))

pheatmap(data[,2:16], 
                    display_numbers = TRUE, 
                    clustering_distance_rows=gower_samples,
                    clustering_distance_cols=gower_features,
                    labels_col = c("Avg Width", "Avg Thickness Backing/Adhesive", "Avg Thickness Backing", "Surface Color", "Surface Feature", "Wide", "Stereomicroscope Color", "Texture", "Bubbles", "Particle", "Edge Texture", "Brightness", "PLM Color", "BOPP", "Angle of Extinction"),
                    labels_row = c("3M 810 Scotch - Tape 1", "3M 600 Scotch - Tape 2", "Scotch Double-Sided - Tape 3", "Scotch Gift-Wrap - Tape 4", "Scotch Transparent - Tape 5", "Scotch Super-Hold - Tape 6", "Scotch Wall-Safe - Tape 8", "Scotch Scrapbooking - Tape 9", "Scotch Single Sided - Tape 10", "Scotch MultiTask - Tape 11", "Scotch Poster - Tape 12", "Staples Invisible - Tape 13", "Amazon Matte - Tape 14", "Amazon Invisible - Tape 15", "Amazon Transparent - Tape 16", "Pen+Gear Invisible - Tape 17", "Pen+Gear Transparent - Tape 18", "Pen+Gear Double-Sided - Tape 19", "Up & Up Transparent - Tape 20", "Up & Up Invisible - Tape 21", "Up & Up Double-Sided - Tape 22", "Up & Up Gift - Tape 23"),
                    angle_col = 3,
                    fontsize = 8,
                    cellheight = 7.8,
                    cellwidth = 19
                    )

```
## Multidimensional Scaling

Multidimensional scaling provides another method of visualizing these clusters and their distances from each other on a two dimensional plot. Although the data for each tape is multivariate, multidimensional scaling preserves the distances between pairwise observations when plotting. Multidimensional scaling also takes in the distances calculated from @sec-distances and returns a set of 22 points where the distances between each point is approximately equal to the dissimilarities between each point. These points are then plotted on a two dimensional scatter plot to visualize the "closeness" of the tapes and can be viewed below. 

```{r}
#| echo: false
#| warning: false


#function that calculates the distance between two tapes

pair_distance <- function(first, second){
  sum <- 0
  for (i in 2:4){
    val <- (first[i] - second[i])^2
    sum <- val + sum
  }
  sum <- sum + sum(first[5:16]!=second[5:16])
  return(sum)
}

```

```{r}
#| echo: false
#| warning: false


#unique combos of rows
combos <- combn(seq(1:22), 2)
#empty list for distances of all possible combos of tapes
distances <- c()
#calculates the distance between all possible combos of tapes
for (i in 1:dim(combos)[2]){
  first <- data[combos[,i][1],]
  second <- data[combos[,i][2],]
  distances <- c(distances, pair_distance(first, second))
}

combo_distances <- cbind(t(combos), matrix(distances))
```

```{r, out.width = "200%"}
#| label: fig-mds
#| fig-cap: "Multidimensional Scaling Plot of 22 Different Tapes. Tape 7 is omitted."
#| echo: false
#| warning: false


#multidimensional scaling

#symmetric matrix of our calculated distances

mat <- matrix(0, nrow = 22, ncol = 22)
mat[lower.tri(mat, diag = FALSE)] <- unlist(combo_distances[,3])
mat <- as.dist(mat)


mds <- mat %>% 
  cmdscale() %>%
  as_tibble()

mds <- as.data.frame(mds)
row.names(mds) <- c(seq(1, 6), seq(8, 23))
mds$clust <- c()

colnames(mds) <- c("Dimension 1", "Dimension 2")
ggscatter(mds, x = "Dimension 1", y = "Dimension 2", 
          label = rownames(mds),
          size = 1,
          repel = TRUE, 
          title = "Multidimensional Scaling Plot") +
  theme_bw() +
  theme(axis.title = element_text(size = 14),
        plot.title = element_text(size = 16))
```


# Analysis and Conclusion



